manuscript_figures

### 2024-06-21 scripts/manuscript_figures.Rmd
updated rugosity models
made proportion change plots

# edited control transcet name
mean_rugosity$TRANSECT[mean_rugosity$TRANSECT == "Control 2"] <- "2"

@@

>>>>

@@

Production to biomass ratios (P:B) for each fish species
```{r}

Proportional change per year in kelp cover on y-axis, year on x-axis (by site; one graph for KB and one graph for FW)
```{r}
# use kelp filt dataframe that was filtered from quad data for sugar kelp
# Convert COVER to percent and make new dataaframe
prop.dat <- kelp.filt %>% mutate(COVER = COVER * 100)

# Function to calculate proportional change and standard error per year
calculate_proportional_change <- function(data) {
  data <- data %>%
    # First, aggregate data by SITE and YEAR
    group_by(SITE, YEAR) %>%
    summarise(COVER = mean(COVER, na.rm = TRUE), .groups = "drop") %>%

    # Then calculate proportional change
    arrange(SITE, YEAR) %>%
    group_by(SITE) %>%
    mutate(Proportional_Change = ifelse(lag(COVER) == 0, NA, (COVER - lag(COVER)) / lag(COVER))) %>%

    # Calculate standard error for each site
    mutate(Proportional_Change_SE = sd(Proportional_Change, na.rm = TRUE) / sqrt(sum(!is.na(Proportional_Change)))) %>%
    ungroup()

  return(data)
}

# Calculate proportional change for each site
prop.dat <- calculate_proportional_change(prop.dat)

# Filter the data for each site
data_kb <- prop.dat %>% filter(SITE == "King's Beach")
data_fw <- prop.dat %>% filter(SITE == "Fort Wetherill")

# function to plot
plot_proportional_change <- function(data, site_name, x_limits = c(-2, 20), y_limits = c(0, 30)) {
  ggplot(data, aes(x = Proportional_Change, y = COVER)) +
    geom_point() +
    geom_errorbarh(aes(xmin = Proportional_Change - Proportional_Change_SE, 
                       xmax = Proportional_Change + Proportional_Change_SE),
                  height = 0) +
    # geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Add trend line
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_text(aes(label = YEAR), vjust = -0.5, size = 3) +
    scale_x_continuous(limits = x_limits) +
    scale_y_continuous(limits = y_limits) +
    labs(title = paste("Proportional Change in Kelp Cover at", site_name),
         x = expression("Proportional change " ~ yr^-1),
         y = "Mean Percent Cover (%)") +
    theme_bw() +
    theme(plot.title = element_text(hjust = 0.5))
}

# Assuming data_kb and data_fw are already filtered datasets for each site
data_kb <- prop.dat %>% filter(SITE == "King's Beach")
data_fw <- prop.dat %>% filter(SITE == "Fort Wetherill")

# Plot for King's Beach with customized settings
plot_kb <- plot_proportional_change(data_kb, site_name = "King's Beach", x_limits = c(-1.2, 1.2), y_limits = c(0, 10))
print(plot_kb)

# Plot for Fort Wetherill with customized settings
plot_fw <- plot_proportional_change(data_fw, site_name = "Fort Wetherill", x_limits = c(-0.6, 0.6), y_limits = c(7, 12))
print(plot_fw)

# Save the plots as PNG files
ggsave("output/figures/Kings_Beach_proportional_change.png", plot = plot_kb, width = 10, height = 6, dpi = 300)
ggsave("output/figures/Fort_Wetherill_proportional_change.png", plot = plot_fw, width = 10, height = 6, dpi = 300)
